---
title: "p8105_hw1_amz2148"
output: github_document
date: "2022-09-24"
---

# Problem 1

First we load the `penguins` dataset.

```{r dataset_loading} 
knitr::opts_chunk$set(echo = TRUE) #shows all code chunks
library(tidyverse) #loads tidyverse package
library(plyr) #loads plyr package
data("penguins", package = "palmerpenguins") #loads dataset
```

In the `penguins` dataset, the names of key variables are ``r names(penguins)``. As indicated by the `summary(penguins)` function run below, in this dataset there are 3 categories of species: ``r count(penguins$species)[1:3,1]`` with ``r count(penguins$species)[1:3,2]`` penguins each respectively; there are 3 categories of island: ``r count(penguins$island)[1:3,1]`` with ``r count(penguins$island)[1:3,2]`` penguins respectively; and there are 2 categories of sex: ``r count(penguins$sex)[1:3,1]`` with ``r count(penguins$sex)[1:3,2]`` penguins respectively. Mean bill length is ``r round(mean(penguins$bill_length_mm, na.rm = TRUE), 2)``mm, mean bill depth is ``r round(mean(penguins$bill_depth_mm, na.rm = TRUE), 2)``mm, mean flipper length is ``r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)``mm, and mean body mass is ``r round(mean(penguins$body_mass_g, na.rm = TRUE), 0)``g. Years range from ``r round(min(penguins$year, na.rm = TRUE), 0)`` to ``r round(max(penguins$year, na.rm = TRUE), 0)``, with a mean year of ``r round(mean(penguins$year, na.rm = TRUE), 0)``. The dataset has ``r nrow(penguins)`` rows (observations) and ``r ncol(penguins)`` columns (variables).

```{r summary_stats}
summary(penguins) #provides summary of variable names and values
```

Below is a scatterplot of flipper length in mm (y) vs. bill length in mm (x), with observations colorcoded by species. This scatterplot is created using the `ggplot()` function through the `tidyverse` package.

```{r scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm,col=species)) + geom_point() #creates colorcoded scatterplot
```

Finally, we set the working directory and export the scatterplot above.

```{r export}
setwd("~/Documents/Columbia University/Fall 2022/Data Science/Homeworks/Homework1/p8105_hw1_amz2148") #set working directory
ggsave("p8105_hw1p1_amz2148.pdf", height = 4, width = 6) #export scatterplot
```

# Problem 2

First we create a data frame with various specifications:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r dataframe}
p2_df = tibble(
  random_norm = rnorm(10),
  random_pos = random_norm > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("small", "small", "small", "medium", "medium", "medium", "large", "large", "large", "small"))
)
```

When attempting to take the mean of these variables using the ``pull`` and ``mean`` functions, only the random sample and logical variables result in numeric means (these two are seen below). The other two variables are not numeric and therefore means cannot be calculated. 

The mean of the `random_norm` variable is ``r mean(pull(p2_df, random_norm))``.
The mean of the `random_pos` logical variable is ``r mean(pull(p2_df, random_pos))``.

Below is a code chunk that applies the ``as.numeric`` function to the logical, character, and factor variables. 

```{r as_numeric, eval = FALSE}
as.numeric(p2_df$random_pos)
as.numeric(p2_df$vec_char)
as.numeric(p2_df$vec_factor)
```

When running this code, the logical and factor variables can be treated as numeric variables, which allow one to calculate a mean. But the character variable cannot be treated as a numeric variable, which explains why a mean can still not be calculated even with the ``as.numeric`` function.



